---
title: "p8105_hw3_jdr2191"
author: "Jess Rosenberg"
due date: "14OCT2021"
output: github_document
---

```{r load_libraries}
library(tidyverse)
library(p8105.datasets)
library(lubridate)
library(viridis)
```

# Problem 1

```{r exploratory_analyses, message=FALSE}
data("instacart")

instacartreorders_df = instacart %>%
  mutate(dow = factor(order_dow, ordered = TRUE, levels = c(0,1,2,3,4,5,6))) %>%
  mutate(order_dow2 = order_dow + 1) %>%
  mutate(day_of_week = wday(order_dow2, label = TRUE)) %>%
  mutate(reorder2 = ifelse(reordered %in% c(0), "No", "Yes") %>%
           factor(levels = c("No","Yes")))

instacartreorders_df %>%
  sample_n(500) %>%
  ggplot(aes(x = order_hour_of_day, fill = reorder2)) +
  geom_density(aes(fill = reorder2), alpha = .5) +
  facet_grid(rows = vars(day_of_week)) +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5)) +
  labs(x = "Hour ordered", y = "Density") +
  guides(fill = guide_legend(title = "Has the item been ordered before?")) +
  ggtitle("The sum of items ordered in each aisle among aisles with 
          more than 10,000 items ordered") +
  scale_fill_viridis_d(option = "C")

instacartdepts_df = instacart %>%
  group_by(department_id, department) %>%
  summarise(number_aisle = n_distinct(aisle))
  
instacartdepts_df %>%
  ggplot(aes(x = reorder(department, number_aisle), y = number_aisle)) + 
  geom_bar(stat = "identity", fill = "#DA4F84") +
  coord_flip() +
  labs(x = "Department Name", y = "Number of Aisles") +
  ggtitle("Number of Aisles per Department") +
  theme(plot.title = element_text(hjust = 0.5))

instacartdow_df = instacart %>%
  mutate(dow = factor(order_dow, ordered = TRUE, levels = c(0,1,2,3,4,5,6))) %>%
  mutate(order_dow2 = order_dow + 1) %>%
  mutate(day_of_week = wday(order_dow2, label = TRUE))

instacartdow_df %>%
  group_by(day_of_week) %>%
  summarize(ordersum = n_distinct(order_id)) %>%
  ggplot(aes(x = day_of_week, y = ordersum)) +
  geom_bar(stat = "identity", fill = "#2EB270") +
  geom_text(aes(label = ordersum), position = position_dodge(width = 0.9), 
            vjust = -0.25, size = 4) +
  labs(x = "Day of the Week", y = "Number of Orders Placed") +
  ggtitle("Total number of orders placed by day of the week") +
  theme(plot.title = element_text(hjust = 0.5))
```

The plot "The sum of items ordered in each aisle among aisles with more than 10,000 items ordered" shows us in a sample of 500 observations the density of items that have been ordered before vs. items that have not been ordered before, broken out by day of the week and hour of the day. We can see from this plot that patterns differ depending on the day of the week and that generally speaking, most items are ordered between the hours of 7am and 7pm.  

The plot "Number of Aisles per Department" displays all 21 departments and the number of aisles per each department in descending order. This shows us that the top 3 departments with the most aisles are "personal care", "pantry" and "snacks". 

The plot "Total number of orders placed by day of the week" displays the total number of orders placed by the day of the week. From this table, we see that the most orders are placed on Sundays, followed by Mondays and Saturdays. The fewest orders are placed on Wednesdays.

There are `r count(instacart)` observations and `r ncol(instacart)` variables in the instacart dataset. Each row in the dataset represents an item ordered by a particular customer. Example products ordered include `r head(unique(pull(instacart, product_name)), 5)`. Products come from different aisles, such as `r head(unique(pull(instacart, aisle)), 5)` and different departments, including for example, `r head(unique(pull(instacart, department)), 5)`. The dataset provides detailed information about the hour and day of the week that orders are placed and even tells us the order in which products were added to orders.  

**How many aisles are there, and which aisles are the most items ordered from?**

```{r aisles, message=FALSE}
aisles = 
  instacart %>% 
    group_by(aisle_id, aisle) %>% 
    summarize(sum_products_ordered = n()) %>% 
    arrange(desc(sum_products_ordered))

view(aisles)
```
There are `r nrow(aisles)` aisles. The top four aisles for the most items ordered from are `r head(pull(aisles, aisle), 4)`.

**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**

```{r aisles_plot, message=FALSE}
aislesplot = 
    instacart %>%
    group_by(aisle_id, aisle) %>%
    summarize(sum_products_ordered = n()) %>%
    filter(sum_products_ordered > 10000)

aislesplot %>%
  ggplot(aes(x = reorder(aisle, sum_products_ordered), 
             y = sum_products_ordered)) + 
  coord_flip() +
  geom_bar(stat = "identity", fill = "#6699ff") +
  ggtitle("The sum of items ordered in each aisle among aisles 
          with more than 10,000 items ordered") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Aisle Name") + ylab("Total number of products ordered")
```

**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r table_threemostpopular, message=FALSE}
topthree =
  instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle_id, aisle) %>%
  count(product_name, name = "number_ordered") %>%
  arrange(desc(number_ordered),.by_group = TRUE) %>%
  slice_head(n = 3) %>%
  subset(select = -c(aisle_id))

knitr::kable(topthree)
```

**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**

```{r pinkladdy_coffeeicecream, message=FALSE}
appleicecream =
  instacart %>%
  mutate(order_dow2 = order_dow + 1) %>%
  mutate(day_of_week = wday(order_dow2, label = TRUE)) %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  subset(select = c(day_of_week, order_hour_of_day, product_name)) %>%
  group_by(day_of_week, product_name) %>%
  mutate(meanhour = round(mean(order_hour_of_day)), 0) %>%
  subset(select = -c(order_hour_of_day)) %>%
  distinct() %>%
  arrange(day_of_week) %>%
  pivot_wider(product_name,
              names_from = "day_of_week",
              values_from = "meanhour"
  )

knitr::kable(appleicecream)
```

# Problem 2

**Data Cleaning Steps:**

* Format the data to use appropriate variable names

* Focus on the “Overall Health” topic

* Include only responses from “Excellent” to “Poor”

* Organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r load_tidy_BRFSS, message=FALSE}
data("brfss_smart2010")

brfss_tidy = brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic %in% "Overall Health") %>%
  filter(response %in% c("Poor","Fair","Good","Very good","Excellent")) %>%
  mutate(response_factor = factor(response, ordered = TRUE, 
                           levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))

view(brfss_tidy)
```

**In 2002, which states were observed at 7 or more locations? What about in 2010?**

```{r states_2002_2010, message=FALSE}
brfss_2002locations = brfss_tidy %>%
  filter(year == 2002) %>% 
  subset(select = c(year,locationabbr, locationdesc)) %>%
  unique() %>%
  group_by(locationabbr) %>%
  summarize(locationdesc = n()) %>%
  filter(locationdesc >= 7)

view(brfss_2002locations)

brfss_2010locations = brfss_tidy %>%
  filter(year == 2010) %>% 
  subset(select = c(year,locationabbr, locationdesc)) %>%
  unique() %>%
  group_by(locationabbr) %>%
  summarize(locationdesc = n()) %>%
  filter(locationdesc >= 7)

view(brfss_2010locations)
```
In 2002, 6 states were observed at 7 or more locations. They are `r unique(pull(brfss_2002locations, locationabbr))`.

In 2010, 14 states were observed at 7 or more locations. They are `r unique(pull(brfss_2010locations, locationabbr))`.

**Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state.**

```{r avg_value_plot, message=FALSE}
brfssexcellent_df = brfss_tidy %>%
  filter(response == "Excellent") %>%
  subset(select = c(year, locationabbr, locationdesc, data_value)) %>%
  unique() %>%
  na.omit() %>%
  subset(select = c(year, locationabbr, data_value)) %>%
  group_by(locationabbr, year) %>%
  mutate(meandatavalue = mean(data_value)) 

brfssexcellent_df2 = brfssexcellent_df %>%
  subset(select = c(year, locationabbr,meandatavalue)) %>%
  unique()

view(brfssexcellent_df2)

brfssexcellent_df2 %>%
  ggplot(aes(x = year, y = meandatavalue)) + 
  geom_line(aes(group = locationabbr, color = locationabbr), size = .9, alpha = .5) +
  geom_point(aes(group = locationabbr, color = locationabbr), size = .4, alpha = .5) +
  labs(x = "Year", y = "Mean data value", col = "State") +
  ggtitle("Among excellent responses, the average data value by state") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

**Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.**

```{r 2006_2010_plots, message=FALSE}
brfss20062010_df = brfss_tidy %>%
  filter(year %in% c(2006, 2010)) %>%
  filter(locationabbr == "NY")

brfss20062010_df %>%
  mutate(year2 = factor(year, levels = c("2006", "2010"))) %>%
  ggplot(aes(x = data_value, fill = response_factor)) + 
  geom_density(aes(fill = response), alpha = .5) + 
  facet_grid(rows = vars(year2), switch = "y") +
  theme_bw() + 
  theme(legend.position = "bottom", axis.title.y = element_blank()) +
  labs(x = "Data Value", fill = "Response") +
  scale_fill_viridis_d(option = "A") +
  ggtitle("The distribution of data value for responses among locations in NY state") +
  theme(plot.title = element_text(hjust = 0.5)) 
  
view(brfss20062010_df)
```

# Problem 3

**Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).**

```{r load_tidy_accel, message=FALSE}
accel = read_csv("accel_data.csv") %>%
        janitor::clean_names() %>%
        mutate(day = factor(day, ordered = TRUE,
                           levels = c("Monday", "Tuesday", "Wednesday", "Thursday", 
                                      "Friday", "Saturday", "Sunday")),
               week = factor(week, ordered = TRUE,levels = c(1:5)),
               day_class = ifelse(day %in% c("Saturday", "Sunday"),"weekend", "weekday") %>%
               factor(levels = c("weekday", "weekend"))) %>%
        rename_at(vars(matches("activity_")), ~str_remove(., "activity_")) %>%
        rename_with(.cols = "1":"1440", function(x){paste0("minute_", x)}) %>%
        pivot_longer(
          cols = starts_with("minute"),
          names_to = "minute",
          names_prefix = "minute_",
          values_to = "activity",
          values_drop_na = TRUE
        ) %>%
        mutate(minute = as.numeric(minute)) %>%
        rename(activity_count = activity)
```
There are `r count(accel)` observations and `r ncol(accel)` variables in the tidied accelerometer dataset. The dataset includes data over 35 days (5 weeks). The dataset also indicates the day of the week that the measurements come from and an activity count for every minute in a 24-hour period. Lastly, there is a newly added variable day_class that indicates if the day of the week is a weekday or a weekend. 

**Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?**

```{r aggregate_accel_trends, message=FALSE}
accelagg = accel %>%
  group_by(day_id, day) %>%
  summarize(agg_sum = sum(activity_count))

knitr::kable(accelagg)
view(accelagg)

accelagg %>%
  ggplot(aes(x = day, y = agg_sum, group = day), color = factor(day)) + 
  geom_boxplot(aes(fill = day)) +
  labs(x = "Day of the week", y = "Total activity", 
       col = "Day of the week") +
  ggtitle("Total activity by day of the week") +
  theme(plot.title = element_text(hjust = 0.5)) 
```


After I made the table, I plotted these data in a boxplots in order to better observe trends. Using the boxplots, we see that the patient is most active on Fridays. His activity is highly variable on Saturdays and not very variable on Wednesdays. 

**Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.**

```{r accelermoeter_plot, message=FALSE}
accel %>%
  ggplot(aes(x = minute, y = activity_count, group = day, color = day)) + 
  geom_point(size = .4, alpha = .5) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), size = 0.4, alpha = 0.5) +
  labs(x = "Minute", y = "Total accelerometer activity", col = "Day of week") +
  ggtitle("24-hour activity time courses per day and by day of the week") +
  theme(plot.title = element_text(hjust = 0.5)) 
  
```

